---
title: "COVID-19 Impacts More on Women and Develops More Seriously on the Old"
author: "Tianxiao Ma"
date: 2020-09-27T21:13:14-05:00
bibliography: 'references.bib'
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
# install.packages("opendatatoronto")
# install.packages("devtools")
# install.packages("tidyverse")
# devtools::install_github("sharlagelfand/opendatatoronto")
# install.packages(knitr)
library(knitr)
library(opendatatoronto)
library(tidyverse)
library(ggplot2)
Covid <- search_packages("COVID")
res <- list_package_resources("64b54586-6180-4485-83eb-81e8fae3b8fe")
get_resource(resource = res[1, "id"])
res[["id"]]
COVID_cases_original <- get_resource("e5bf35bc-e681-43da-b2ce-0242d00922ad")
```
```
```
# Abstract

  With the COVID-19 cases investigation data released by the Toronto Public
  Health, it turns out that over half of the confirmed and probable cases in Toronto
  are composed by female patients, and patients over 50-year-old have a higher ICU
  admission rate.
  @citetidyverse

# Introduction

  In order to analyze the impact of COVID-19 on our society, the Toronto Public Health has been collecting COVID-19 cases data in Toronto since the first confirmed COVID-19 case in January, 2020. We want to know which group is been affected the most as COVID-19 continues to threaten more lives. In a total of 17872 cases, 52.7% of them are composed by female patients, and 45.8% of them are male patients. From the data, we observed that COVID-19 has an bigger impact on Toronto female residents than male residents. Additionally, the impact of COVID-19 to an individual also differs across different age groups. If the level of impact is defined by whether the patient has been to ICU, we can observe that patients over 50 years old are more likely to develop serious illness, and patients with the age group of 70 to 79 years old have the highest ICU admission rate. In conclusion, COVID-19 has the most impact on females and develops serious among people over 50 years old.
  
# About the Data
  The dataset I used is downloaded through the Toronto Open Data Portal which contains
all the confirmed and probable COVID-19 cases managed by the Toronto Public Health. The latest update by the time of this project is on 23, September, 2020.
The dataset is been refreshed and overwritten every Wednesday. It forms as a table which contains 18 variables related to demographic, geographic, and severity information. Each row represents an 
independent patient with all the associated information. 
  The database contains 17872 rows in total until Sep 23, 2020.
  CITATION: Toronto Public Health, Sept 23, 2020. ‘COVID-19 CASES IN TORONTO’ https://open.toronto.ca/dataset/covid-19-cases-in-toronto/

  Client Gender is the self-reported gender and generally based on people's assigned biological sex.

  Age Group is the patient's age at the time of illness , in years. This variable contains 10 elements which are 19 and younger, 20 to 29 Years, 30 to 39 Years, 40 to 49 Years, 50 to 59 Years, 60 to 69 Years, 70 to 79 Years, 80 to 89 Years, 90 and older, and unknown(blank).

  Ever in ICU indicates whether if cases were admitted to the intensive care unit (ICU) 
related to their COVID-19 infection, it includes cases are currently in ICU and those been discharged or deceased. 

# About my clean vision data  
```{r data-prepare}
clean_data <- COVID_cases_original %>%
  select(`Age Group`, `Client Gender`,`Ever in ICU`)
head(clean_data)
# I create a table only containing four variables I want to analyze, which are 
# Age Group, Client Gender, and Ever in ICU.
```
  I selected 3 variables to create a new table from the original dataset and 
named the new table clean_data. Since I want to investigate the demographic and
severity information, clean_data contains three variables that are 
Client Gender, Age Group and Ever in ICU. 

# Visualize gender variable in a bar plot
```{r}
gender_proportion <- clean_data %>%
  ggplot(aes(x = `Client Gender`))+
  geom_bar() +
  labs(caption ='Toronto Public Health, Sept 23, 2020. COVID-19 CASES IN TORONTO')+ 
  theme_light()
gender_proportion
```

# 1st Summary Statistics based on Client Gender
  Based on the statistics, We observed that the number of female patients is the largest, which is 9427 out of 17872 patients, while there are only 5 transgender patients. 
```{r}
Gender_sum <- clean_data %>%
  group_by(`Client Gender`) %>% 
    count()
Gender_sum
# Count number of each element in Client gender.
```
  
## Discuission -- Why female? 
  From the data collected from 17872 Toronto COVID patients, we can observe that females are being affected the most by COVID-19. There's an observable difference between the number of male and female patients which is composed by 45.8% and 52.7% of the total cases population. One potential reason for higher infection rate is 
that females tend to have a higher participation rate working in service industries(Kwolek-Folland, A., & Walsh, M. 2007), such as restaurents and healthcare facilities etc.
  Citation: Kwolek-Folland, A., & Walsh, M. (2007). Women in the Service Industries: National Perspectives. The Business History Review, 81(3), 421-427. Retrieved September 27, 2020, from http://www.jstor.org/stable/25097375

# 2nd Graph -- Scatterplot between age group and ICU proportion rate in each group
```{r}
new <- clean_data %>%
  group_by(`Age Group`) %>%
  filter(`Age Group` != 'NaN') %>%
  summarise(proportion = mean(`Ever in ICU` == 'Yes'),.groups = 'drop')
new
# ICU rate in different age groups without unknown age group.

age_and_icu_scatter <- new %>%
   ggplot(aes(x=`Age Group`, y=proportion)) + geom_point(shape=24, color="red")+
  labs(x= 'Age Group', y = 'ICU proportion',caption = 'Toronto Public Health, Sept 23, 2020. COVID-19 CASES IN TORONTO')+
  coord_flip() +
  ggtitle("Scatterplot of age group and each group's ICU proportion")
age_and_icu_scatter

```

# 2nd Graph discuission based on age group and its icu proportion
  Without the influence of blink data in age group, I expect to see ICU proportion
  and age group have a positive relationship, that is, ICU proportion rate 
  will increase as age increases.The scatterplot describes that people
  with age group 70-79 years has the highest ICU proportion which is around 0.0794,
  and people with age group 20-29 years has the lowest ICU proportion which is 
  around 0.0043. However,people over 79 years old have a low 
  ICU proportion especially for the group 90 and older, with only 0.005%.

# Discuission -- Why COVID hurt 70-79 years old so much? Weakness and potential missing data.
  According to Professor Hassan Vally, he noticed human's immune system weakens 
  as we age. In the graph above I only analyzed if the patient has ever been sent to ICU 
  without the consideration of death rate. We observed the ICU proportion starts 
  to increase for those over 50 years old, then there's a significant decrease 
  for people over 79 years of age. However, the death rate begins to increase for
  people over 50 years old and reaches the peak at over 90 years old age group. 
  The cruel reality is that many elderly people living alone have passed away without
  being noticed at home, they can be accounted to missing data.
  Another thing causing the misleading ICU proportion is for those old people
  who are sent to hospital can't take their last breath before transfering to ICU
  due to their weak immune system.
  
# References
  H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
  Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686
  
  JJ Allaire and Yihui Xie and Jonathan McPherson and Javier Luraschi and Kevin Ushey and Aron Atkins
  and Hadley Wickham and Joe Cheng and Winston Chang and Richard Iannone (2020). rmarkdown: Dynamic
  Documents for R. R package version 2.3. URL https://rmarkdown.rstudio.com.
  
  Yihui Xie (2020). knitr: A General-Purpose Package for Dynamic Report Generation in R. R package
  version 1.30.
  
  Toronto Public Health, Sept 23, 2020. 'COVID-19 CASES IN TORONTO'
  https://open.toronto.ca/dataset/covid-19-cases-in-toronto/
  
  Kwolek-Folland, A., & Walsh, M. (2007). Women in the Service Industries: 
  National Perspectives. The Business History Review, 81(3), 421-427. 
  Retrieved September 27, 2020, from http://www.jstor.org/stable/25097375